---
title: "Messi Viz Project"
author: "Christian Azumah"
date: "2023-02-13"
output: html_document
---

#Importing the Data

Step 1 - Install the packages that will be used.

```{r}
install.packages("devtools")
install.packages("remotes")
```

```{r}
install.packages("cli")
```

```{r}
install.packages("purrr")
```

```{r}
install.packages("extrafont")
```


Step 1a - Use 'remote package' to install all the tools to be used. 

```{r}
remotes::install_version("SDMTools", "1.1-221")
```

Step 1b - Install StatsBombR Package. 

```{r}
devtools::install_github("statsbomb/StatsBombR")
```

```{r}
my_packages <- library()$results[,1]
head(my_packages, 10)
```

Step 2 - Load the Packages which will be used. 

```{r}
library(tidyverse)
library(devtools)
library(ggplot2)
library(StatsBombR)
```

##Importing the Data and Creating a DataFrame
 
```{r}
FAWSL <- FreeCompetitions()%>%
  filter(competition_id==37 & season_name=="2020/2021")
```

```{r}
Matches <- FreeMatches(FAWSL)
```

```{r}
StatsBombData_FAWSL <- free_allevents(MatchesDF = Matches,Parallel = T)
```

```{r}
StatsBombData_FAWSL = allclean(StatsBombData_FAWSL)
```

```{r}
names(StatsBombData_FAWSL)
```

#Data Exploration (Goals & Shots)

This next section is going to explore the Data Use Cases involving Goals & Shots.

##Extracting Goals & Shots

The first thing to do would be to group the data by team, so that operations are handled on a team by team basis. The summarise function then takes that operation and produces a new separate table out of it.

```{r}
shots_goals = StatsBombData_FAWSL %>%
  group_by(team.name) %>%
  summarise(shots = sum(type.name == "Shot", na.rm = TRUE),
            goals = sum(shot.outcome.name == "Goal", na.rm = TRUE))
```

The table (shots_goals) produced above shows the shots and goals total for each team. 

To investigate on a per game basis the code above can be modified using n_distinct from the dplyr package. 

```{r}
shots_goals_pg = StatsBombData_FAWSL %>%
  group_by(team.name) %>%
  summarise(shots = sum(type.name == "Shot", na.rm = TRUE)/n_distinct(match_id),
            goals = sum(shot.outcome.name == "Goal", na.rm = TRUE)/n_distinct(match_id))
```

##Transformation into a Chart

```{r}
ggplot(data = shots_goals_pg,
       aes(x = reorder(team.name,shots), y = shots)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(y = "Shots") +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip() +
  theme_SB()
```

##Player Shots per 90

```{r}
player_shots = StatsBombData_FAWSL %>%
  group_by(player.name, player.id) %>%
  summarise(shots = sum(type.name == "Shot", na.rm = TRUE))
```

Here we've done the Same calculation as above, done with regards to Player Shots. Included player.id as will be needed later. 

```{r}
player_minutes = get.minutesplayed(StatsBombData_FAWSL)
```

```{r}
player_minutes = player_minutes %>%
  group_by(player.id) %>%
  summarise(minutes = sum(MinutesPlayed))
```

Above we get the player minutes and summarize the new outputs. 

```{r}
player_shots = left_join(player_shots, player_minutes)
```
Next we join Player Shots with Player minutes.

```{r}
player_shots = player_shots %>% mutate(nineties = minutes/90)
```

Then 

```{r}
player_shots = player_shots %>% mutate(shots_per90 = shots/nineties)
```

And Finally. 

##Plotting Passes

```{r}
devtools::install_github("FCrSTATS/SBpitch")
```

```{r}
library(SBpitch)
```

Next we are going to extract passes from Viv. 

```{r}
wsldata <- free_allevents(MatchesDF = Matches,Parallel = T)
```

```{r}
library(SBpitch)

passes = wsldata %>% 
  filter(type.name=="Pass" & is.na(pass.outcome.name) & 
           player.id==15623) %>%
  filter(pass.end_location.x >= 102 & pass.end_location.y<=62 &
pass.end_location.y>=18)

create_Pitch() +
  geom_segment(data = passes,aes(x=location.x,y=location.y,
                                 xend=pass.end_location.x, yend=pass.end_location.y),
               lineend = "round",size=0.5,colour="#000000", arrow = 
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Vivianne Miedema, Completed Box Passes", subtitle = "WSL,2020-21") +
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)
```

```{r}
passes = wsldata %>% 
  filter(type.name=="Pass" & is.na(pass.outcome.name) & 
           player.id==15623)
```

```{r}
filter(passes, pass.end.location.x >= 102 & pass.end_location.y <= 62 & pass.end_location.y >= 18)
```








```{r}
library(SBpitch) 

passes = wsldata %>%
  filter(type.name=="Pass" & !is.na(pass.outcome.name) &
           player.id==4641) %>%
  filter(pass.end_location.x >= 102 & pass.end_location.y <= 62 &
           pass.end_location.y >= 18) 

create_Pitch() +
  geom_segment(data = passes, aes(x = location.x, y = location.y,
                                  xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Fran Kirby, Completed Box Passes", subtitle = "WSL,
       2020-21") +
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)
```








